---
title: "Analysis of Longevity"
author: "Ryan Abdella"
date: "January 15, 2015"
output:
  html_document:
    css: ~/Github/Food-Optimization/Scripts/foghorn_edited.css
---


```{r, warning = FALSE, message = FALSE, echo = FALSE}

alt <- substr(experimentName, start = 1, stop = nchar(experimentName) - 1)
processedDataPath <- paste("../Data/Processed/", alt, sep = "")

source("./wMT_fxns_alt.R")
source("./survival_fxns.R")
source(paste("./", experimentName, ".R", sep = ""))

paste(experimentName)

```
 
## Raw Activity ##

```{r, warning = FALSE, echo = FALSE}

## Pull in all the files associated with the current experiment  ##
## and create a data frame containing all the raw data.          ##
## Regroup data frame so that it's ordered by strain, replicate, ##
## and day. Add in strain information.                           ##

files <- dir(path = paste("../Data/Raw/", alt, sep = ""), plateID, full.names = TRUE)

nFiles <- length(files)

fileAttributes.df <- data.frame(binSize = vector(length = nFiles),
                      time = vector(length = nFiles),
                      numGroups = vector(length = nFiles),
                      numBins = vector(length = nFiles))

inputList <- list()

## Need to change to using dplyr instead of plyr. ##
for (i in 1:nFiles) {
  inputList[[i]] <- processMicrotrackerReport(files[i])
  fileAttributes.df$binSize[i] = inputList[[i]]$binSize[1]
  fileAttributes.df$time[i] = inputList[[i]]$time[1]
  fileAttributes.df$numGroups[i] = inputList[[i]]$numGroups[1]
  fileAttributes.df$numBins[i] = inputList[[i]]$numBins[1]
}

experimentAttributes.df <- fileAttributes.df %>% summarise_each(funs(n_distinct))

if (experimentAttributes.df$binSize != 1 || experimentAttributes.df$numGroups != 1) {
  message("Not all wMicrotracker files have the same bin size or number of groups.")
}

minBins <- inputList[[1]]$numBins[1]

for (i in 2:length(inputList)) {
  minBins <- min(minBins, inputList[[i]]$numBins)
}

raw.df <- ldply(inputList, generateDataFrameFromFile) %>% filter(time <= minBins * 2) # need to check consistent binSize instead of 2
raw.df$day <- rep(1:(length(files)), each = nStrains * replicates * minBins)

save(raw.df, file = paste(processedDataPath, "/", experimentName, "_Raw2.Rda", sep = ""), ascii = TRUE)

ordered.df <- raw.df[order(raw.df$col, raw.df$row, raw.df$day), ]

ordered.df$strain <- factor(ordered.df$col, labels = strains)

save(ordered.df, file = paste(processedDataPath, "/", experimentName, "_Processed2.Rda", sep = ""), ascii = TRUE)

```

```{r Raw2, warning = FALSE, message = FALSE, echo = FALSE, fig.path = paste("Scripts/", processedDataPath, "/", experimentName, "_", sep = ""), fig.width = 15, fig.height = 12}

ggplot(ordered.df) + aes(x = day, y = activity) + geom_point() + facet_grid(row ~ strain) +
  labs(x = "Time (day)", y = "Raw Activity")

```

## Cleaned Data ##

```{r, warning = FALSE, message = FALSE, echo = FALSE}

## Remove both columns and individual wells that were contaminated. ##

cleaned.df <- ordered.df
#cleaned.df$activity[(ordered.df$col %in% colRemove)] <- NA
#for(i in 1:length(wellRemove$row)) {
  #cleaned.df$activity[cleaned.df$row == wellRemove$row[i] & cleaned.df$col == wellRemove$col[i]] <- NA
#}

save(cleaned.df, file = paste(processedDataPath, "/", experimentName, "_Cleaned2.Rda", sep = ""), ascii = TRUE)

```

```{r Cleaned2, warning = FALSE, message = FALSE, echo = FALSE, fig.path = paste("Scripts/", processedDataPath, "/", experimentName, "_", sep = ""), fig.width = 15, fig.height = 12}

ggplot(cleaned.df) + aes(x = factor(day), y = activity) + geom_boxplot() + facet_grid(row ~ strain) +
  labs(x= "Time (day)", y = "Raw Activity")

```

```{r, warning = FALSE, message = FALSE, echo = FALSE}

well.df <- cleaned.df %>%
  group_by(strain, row, day) %>%
  summarize(mean.activity = mean(activity), 
            sd.activity = sd(activity), 
            median.activity = median(activity, na.rm = TRUE), 
            total.activity = sum(activity))

well.df <- ungroup(well.df)

```

## Well Mean Data ##

```{r Well_Mean2, warning = FALSE, message = FALSE, echo = FALSE, fig.path = paste("Scripts/", processedDataPath, "/", experimentName, "_", sep = ""), fig.width = 15, fig.height = 12}

ggplot(well.df) + geom_line(aes(x = day, y = mean.activity)) + 
  geom_errorbar(aes(x = day, ymin = mean.activity - sd.activity, ymax = mean.activity + sd.activity)) + 
  facet_grid(row ~ strain) + labs(x = "Time (day)", y = "Mean Activity")

```

## Well Median Data ##

```{r Well_Median2, warning = FALSE, message = FALSE, echo = FALSE, fig.path = paste("Scripts/", processedDataPath, "/", experimentName, "_", sep = ""), fig.width = 15, fig.height = 12}

ggplot(well.df) + aes(x = day, y = median.activity) + geom_line() + facet_grid(row ~ strain) +
  labs(x = "Time (day)", y = "Median Activity")

```

## Well Total Data ##

```{r Well_Total2, warning = FALSE, message = FALSE, echo = FALSE, fig.path = paste("Scripts/", processedDataPath, "/", experimentName, "_", sep = ""), fig.width = 15, fig.height = 12}

ggplot(well.df) + aes(x = day, y = total.activity) + geom_line() + facet_grid(row ~ strain) +
  labs(x = "Time (day)", y = "Aggregate Activity")

```

## Well Normalized Data ##

```{r, warning = FALSE, message = FALSE, echo = FALSE}

## Normalize activity data by the maximum activity of each well. ##
## NEED TO ADD NORMALIZING BY THE NUMBER OF WORMS SORTED.        ##

well.df <- well.df %>%
  group_by(strain, row) %>%
  mutate(norm.act = 100 * total.activity / max(total.activity, na.rm = TRUE))

well.df <- ungroup(well.df)

save(well.df, file = paste(processedDataPath, "/", experimentName, "_Well2.Rda", sep = ""), ascii = TRUE)

```

```{r Well_Normalized2, warning = FALSE, message = FALSE, echo = FALSE, fig.path = paste("Scripts/", processedDataPath, "/", experimentName, "_", sep = ""), fig.width = 15, fig.height = 12}

ggplot(well.df) + aes(x = day, y = norm.act) + ylim(0, 100) + geom_line() + facet_grid(row ~ strain) +
  labs(x = "Time (day)", y = "Normalized Activity")

```

## Setting Top of Curve by Well ##

```{r, warning = FALSE, message = FALSE, echo = FALSE}

well.df <- well.df %>%
  group_by(strain, row) %>%
  mutate(max.sum = day[which.max(norm.act)[1]]) %>%
  group_by(row) %>%
  mutate(adjusted.activity = ifelse(day < max.sum, 100, norm.act))
well.df <- ungroup(well.df)
  
```

```{r Well_Adjusted2, warning = FALSE, message = FALSE, echo = FALSE, fig.path = paste("Scripts/", processedDataPath, "/", experimentName, "_", sep = ""), fig.width = 15, fig.height = 12}

ggplot(well.df) + aes(x = day, y = adjusted.activity) + geom_line() + facet_grid(row ~ strain) +
  labs(x = "Time (day)", y = "Adjusted Activity")

```

## Well Curve Fitting ##

```{r, warning = FALSE, message = FALSE, echo = FALSE}

well_params.df <- data.frame(strain = rep(strains, each = length(levels(well.df$row))), row = rep(levels(well.df$row), nStrains), bparam = rep(0, nStrains * replicates), cparam = rep(0, nStrains * replicates))

l = 1

for (x in levels(well.df$strain)) {
  for (y in levels(well.df$row)) {
    if (!(is.na(well.df$norm.act[well.df$row == y & well.df$strain == x]))[1] == TRUE) {
      temp <- nlm(well_twoplog, c(2, 1), well.df, y, x)[["estimate"]]
      well_params.df$bparam[l] <- temp[1]
      well_params.df$cparam[l] <- temp[2]
      }
    l = l + 1
    }
  }

save(well_params.df, file = paste(processedDataPath, "/", experimentName, "_Well_Params2.Rda", sep = ""), ascii = TRUE)

print(well_params.df)

```

## 2 Parameter Logistic Function Fits by Well ##

```{r, warning = FALSE, message = FALSE, echo = FALSE}

times <- 1:length(levels(factor(well.df$day)))

draw_twoplog <- function(bparam, cparam) {
  return((100 / (1 + ((times / cparam) ^ bparam))))
}

well_survivals.df <- data.frame(day = times)

l <- 2

for (x in levels(well.df$strain)) {
  for (y in levels(well.df$row)) {
    well_survivals.df[l] <- draw_twoplog(well_params.df$bparam[well_params.df$strain == x & well_params.df$row == y], well_params.df$cparam[well_params.df$strain == x & well_params.df$row == y])
    l <- l + 1
  }
}

well_survivals.df <- well_survivals.df %>% gather(strain, yvalue, 2:97)
well.df$yvalue <- well_survivals.df$yvalue

for (i in 1:length(well.df$yvalue)) {
  if (is.na(well.df$norm.act[i]) == TRUE) {
    well.df$yvalue[i] <- NA
  }
}

```

```{r Well_Fits2, warning = FALSE, message = FALSE, echo = FALSE, fig.path = paste("Scripts/", processedDataPath, "/", experimentName, "_", sep = ""), fig.width = 15, fig.height = 12}

ggplot(well.df) + geom_line(aes(x = day, y = norm.act)) + ylim(0, 100) + geom_line(aes(x = day, y = yvalue, color = "red")) + facet_grid(row ~ strain)

```

## Strain Mean Data ##

```{r, warning = FALSE, message = FALSE, echo = FALSE}

## Let's look at the mean and standard deviation of the ##
## data per strain.                                     ##

strain.df <- well.df %>%
  group_by(strain, day) %>%
  summarise(mean.activity = mean(total.activity, na.rm = TRUE), 
            sd.activity = sd(total.activity, na.rm = TRUE), 
            median.activity = median(total.activity, na.rm = TRUE), 
            total.activity = sum(total.activity, na.rm = TRUE))

strain.df <- ungroup(strain.df)

for (i in 1:length(strain.df$total.activity)) {
  if (strain.df$total.activity[i] == 0) {
    strain.df$total.activity[i] <- NA
  }
  if (!is.finite(strain.df$mean.activity[i])) {
    strain.df$mean.activity[i] <- NA
  }
  if (!is.finite(strain.df$sd.activity[i])) {
    strain.df$sd.activity[i] <- NA
  }
}

```

```{r Strain_Mean2, warning = FALSE, message = FALSE, echo = FALSE, fig.path = paste("Scripts/", processedDataPath, "/", experimentName, "_", sep = ""), fig.width = 15, fig.height = 12}

ggplot(strain.df) + geom_line(aes(x = day, y = mean.activity)) + 
  geom_errorbar(aes(x = day, ymin = mean.activity - sd.activity, ymax = mean.activity + sd.activity)) + 
  facet_wrap( ~ strain) + labs(x = "Time (day)", y = "Mean Activity")

```

## Strain Median Data ##

```{r Strain_Median2, warning = FALSE, message = FALSE, echo = FALSE, fig.path = paste("Scripts/", processedDataPath, "/", experimentName, "_", sep = ""), fig.width = 15, fig.height = 12}

ggplot(strain.df) + aes(x = day, y = median.activity) + geom_line() + facet_wrap( ~ strain) +
  labs(x = "Time (day)", y = "Median Activity")

```

## Strain Total Data ##

```{r Strain_Total2, warning = FALSE, message = FALSE, echo = FALSE, fig.path = paste("Scripts/", processedDataPath, "/", experimentName, "_", sep = ""), fig.width = 15, fig.height = 12}

ggplot(strain.df) + aes(x = day, y = total.activity) + geom_line() + facet_wrap( ~ strain) +
  labs(x = "Time (day)", y = "Activity")

```

## Strain Normalized Data ##

```{r, warning = FALSE, message = FALSE, echo = FALSE}

strain.df <- strain.df %>%
  group_by(strain) %>%
  mutate(norm.act = 100 * total.activity / max(total.activity, na.rm = TRUE))

strain.df <- ungroup(strain.df)

```

```{r Strain_Normalized2, warning = FALSE, message = FALSE, echo = FALSE, fig.path = paste("Scripts/", processedDataPath, "/", experimentName, "_", sep = ""), fig.width = 15, fig.height = 12}

ggplot(strain.df) + aes(x = day, y = norm.act) + ylim(0, 100) + geom_line() + facet_wrap( ~ strain) + 
  labs(x = "Time (day)", y = "Normalized Activity")

```

## Setting Top of Curve by Strain ##

```{r, warning = FALSE, message = FALSE, echo = FALSE}

strain.df <- strain.df %>%
  group_by(strain) %>%
  mutate(max.sum = day[which.max(norm.act)[1]]) #%>%
  #mutate(adjusted.activity = ifelse(day < max.sum, 100, norm.act))

strain.df <- ungroup(strain.df)
strain.df$adjusted.activity <- rep(0, length(strain.df$day))

for (i in 1:length(strain.df$strain)) {
  strain.df$adjusted.activity[i] <- ifelse(strain.df$day[i] < strain.df$max.sum[i], 100, strain.df$norm.act[i])
}

```

```{r Strain_Adjusted2, warning = FALSE, message = FALSE, echo = FALSE, fig.path = paste("Scripts/", processedDataPath, "/", experimentName, "_", sep = ""), fig.width = 15, fig.height = 12}

#ggplot(strain.df) + aes(x = day, y = adjusted.activity) + geom_line() + facet_grid(row ~ strain) +
#  labs(x = "Time (day)", y = "Adjusted Activity")

```

## Strain Curve Fitting ##

```{r, warning = FALSE, message = FALSE, echo = FALSE}

strain_params.df <- data.frame(strain = strains, bparam = rep(0, nStrains), cparam = rep(0, nStrains))

l = 1

for (x in strains) {
  if(!(is.na(strain.df$norm.act[strain.df$strain == x]))[1] == TRUE) {
    temp <- nlm(strain_twoplog, c(2, 2), strain.df, x)[["estimate"]]
    strain_params.df$bparam[l] <- temp[1]
    strain_params.df$cparam[l] <- temp[2]
  }
  l = l + 1
}

save(strain_params.df, file = paste(processedDataPath, "/", experimentName, "_Strain_Params2.Rda", sep = ""), ascii = TRUE)

print(strain_params.df)

```

## 2 Parameter Logistic Function Fits by Strain ##

```{r, warning = FALSE, message = FALSE, echo = FALSE}

times <- 1:length(levels(factor(strain.df$day)))

strain_survivals.df <- data.frame(day = times)

l <- 2

for (x in levels(strain.df$strain)) {
    strain_survivals.df[l] <- draw_twoplog(strain_params.df$bparam[strain_params.df$strain == x], strain_params.df$cparam[strain_params.df$strain == x])
    l <- l + 1
}

strain_survivals.df <- strain_survivals.df %>% gather(strain, yvalue, 2:13)

strain.df$yvalue <- strain_survivals.df$yvalue

for (i in 1:length(strain.df$yvalue)) {
  if (is.na(strain.df$norm.act[i]) == TRUE) {
    strain.df$yvalue[i] <- NA
  }
}

```

```{r Strain_Fits2, warning = FALSE, message = FALSE, echo = FALSE, fig.path = paste("Scripts/", processedDataPath, "/", experimentName, "_", sep = ""), fig.width = 15, fig.height = 12}

ggplot(strain.df) + geom_line(aes(x = day, y = norm.act)) + ylim(0, 100) + geom_line(aes(x = day, y = yvalue, color = "red")) + facet_wrap( ~ strain)

```

## Implementing NLS ##

```{r, warning = FALSE, message = FALSE, echo = FALSE}

## This should only be necessary if the curve fits to the data are not accurate at all. ##

#nlsSummary.df <- data.frame(day = summary.df$day[summary.df$strain == "CX11314"], norm.act = summary.df$norm.act[summary.df$strain == "CX11314"])
#nlsTest <- nls(norm.act ~ SSlogis(day, Asym, xmid, scal), nlsSummary.df)
#nlm(strain_twoplog, c(1, coef(nlsTest)[["xmid"]]), summary.df, "CX11314")

```
